# example 11.1 of section 11.2.1 
# (example 11.1 of section 11.2.1)  : Documentation and deployment : Using R markdown to produce milestone documentation : What is R markdown? 
# Title: R-annotated Markdown 

---                                         	# Note: 1 
title: "Buzz scoring example"
output: github_document
---

```{r, include = FALSE}                     	# Note: 2 
# process document with knitr or rmarkdown.
# knitr::knit("Buzz_score_example.Rmd") # creates Buzz_score_example.md
# rmarkdown::render("Buzz_score_example.Rmd", 
#                   rmarkdown::html_document()) # creates Buzz_score_example.html
```                                          	# Note: 3 

Example scoring (making predictions with) the Buzz data set.      	# Note: 4 

First attach the `randomForest` package and load the model and test data.

```{r}                                       	# Note: 5 
suppressPackageStartupMessages(library("randomForest"))

lst <- readRDS("thRS500.RDS")
varslist <- lst$varslist
fmodel <- lst$fmodel
buzztest <- lst$buzztest
rm(list = "lst")
```

Now show the quality of our model on held-out test data.            	# Note: 6 

```{r}                                         	# Note: 7 
buzztest$prediction <- predict(fmodel, newdata = buzztest, type = "prob")[, 2, drop = TRUE]

WVPlots::ROCPlot(buzztest, "prediction", 
                 "buzz", 1, 
                 "ROC curve estimating quality of model predictions on held-out data")
```

# Note 1: 
#   YAML (yet another markup language) header specifying some meta-data: title and default output format. 

# Note 2: 
#   An R markdown “start code chunk” 
#   annotation. The “include = FALSE” directive says 
#   the block is not shown in the rendering. 

# Note 3: 
#   End of the R markdown block, all content between the start and end marks is treated as R code and executed. 

# Note 4: 
#   Free Markdown text. 

# Note 5: 
#   Another R code block. In this case we are loading an already produced random Forest model and test data. 

# Note 6: 
#   More free test. 

# Note 7: 
#   Another R code chunk. 

