---
output: github_document
---



00003_informalexample_2.1_of_section_2.1.2.R


```{r 00003_informalexample_2.1_of_section_2.1.2.R }
# informalexample 2.1 of section 2.1.2 
# (informalexample 2.1 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

print(seq_len(25))
# [1]  1  2  3  4  5  6  7  8  9 10 11 12
# [13] 13 14 15 16 17 18 19 20 21 22 23 24
# [25] 25



```




00004_informalexample_2.2_of_section_2.1.2.R


```{r 00004_informalexample_2.2_of_section_2.1.2.R }
# informalexample 2.2 of section 2.1.2 
# (informalexample 2.2 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

example_vector <- c(10, 20, 30)                 	# Note: 1 
example_list <- list(a = 10, b = 20, c = 30)    	# Note: 2 

example_vector[1]                               	# Note: 3 
## [1] 10
example_list[1]
## $a
## [1] 10

example_vector[[2]]                             	# Note: 4 
## [1] 20
example_list[[2]]
## [1] 20

example_vector[c(FALSE, TRUE, TRUE)]            	# Note: 5 
## [1] 20 30
example_list[c(FALSE, TRUE, TRUE)]
## $b
## [1] 20
## 
## $c
## [1] 30

example_list$b                                  	# Note: 6 
## [1] 20

example_list[["b"]]
## [1] 20

# Note 1: 
#   Build an example vector. c() is R’s concatenate operator—it builds longer vectors and lists 
#   from shorter ones without nesting. For example 
#   c(1) is just the number 1, and c(1, c(2, 3)) is 
#   equivalent to c(1, 2, 3), which in turn is the 
#   integers 1 through 3 (though stored in a floating 
#   point format). 

# Note 2: 
#   Build an example list. 

# Note 3: 
#   Demonstrate vector and list use of []. Notice for the list [] returns a new short list, not the item. 

# Note 4: 
#   Demonstrate vector and list use of [[]]. In common cases [[]] forces a single item to be returned, though for nested lists of complex type this item itself could be a list. 

# Note 5: 
#   Vectors and lists be indexed by vectors of logicals, integers, and (if the vector or list has names) characters. 

# Note 6: 
#   For named examples, the syntax example_list$b is essentially a short-hand for 
#   example_list[["b"]] (the same is true for named 
#   vectors). 



```




00005_informalexample_2.3_of_section_2.1.2.R


```{r 00005_informalexample_2.3_of_section_2.1.2.R }
# informalexample 2.3 of section 2.1.2 
# (informalexample 2.3 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

x <- 1:5
print(x)                                  	# Note: 1 
# [1] 1 2 3 4 5

# x <- cumsumMISSPELLED(x)                  	# Note: 2 
# Error in cumsumMISSPELLED(x) : could not find function "cumsumMISSPELLED"

print(x)                                  	# Note: 3 
# [1] 1 2 3 4 5

x <- cumsum(x)                            	# Note: 4 
print(x)
# [1]  1  3  6 10 15

# Note 1: 
#   Define a value we are interested in and store it in the variable x. 

# Note 2: 
#   Attempt, and fail, to assign a new result to x. 

# Note 3: 
#   Notice in addition to supply a useful error message, R preserved the original value of x. 

# Note 4: 
#   Try the operation again, using the correct spelling of cumsum(). cumsum(), short for cumulative sum, is a useful function that computes running totals quickly. 



```




00006_informalexample_2.4_of_section_2.1.2.R


```{r 00006_informalexample_2.4_of_section_2.1.2.R }
# informalexample 2.4 of section 2.1.2 
# (informalexample 2.4 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

nchar("a string")
# [1] 8

nchar(c("a", "aa", "aaa", "aaaa"))
# [1] 1 2 3 4



```




00007_informalexample_2.5_of_section_2.1.2.R


```{r 00007_informalexample_2.5_of_section_2.1.2.R }
# informalexample 2.5 of section 2.1.2 
# (informalexample 2.5 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

1 +
  2



```




00008_informalexample_2.6_of_section_2.1.2.R


```{r 00008_informalexample_2.6_of_section_2.1.2.R }
# informalexample 2.6 of section 2.1.2 
# (informalexample 2.6 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

1
  + 2



```




00009_informalexample_2.7_of_section_2.1.2.R


```{r 00009_informalexample_2.7_of_section_2.1.2.R }
# informalexample 2.7 of section 2.1.2 
# (informalexample 2.7 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

d <- data.frame(x = c(1, NA, 3))       	# Note: 1 
print(d)
#    x
# 1  1
# 2 NA
# 3  3                                 	# Note: 2 

d$x[is.na(d$x)] <- 0                   	# Note: 3 
print(d)
#   x
# 1 1
# 2 0
# 3 3

# Note 1: 
#   “data.frame” is R’s tabular data type, and the most important data type in R. A data.frame 
#   holds data organized in rows and columns. 

# Note 2: 
#   When printing data.frames, row numbers are shown in the fist (unnamed) column, and column 
#   values are shown under their matching column 
#   names. 

# Note 3: 
#   We can place a slice or selection of the x column of d on the left-hand side of an assignment to easily replace all NA values with zero. 



```




00010_informalexample_2.8_of_section_2.1.2.R


```{r 00010_informalexample_2.8_of_section_2.1.2.R }
# informalexample 2.8 of section 2.1.2 
# (informalexample 2.8 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

d <- data.frame(x = 1, y = 2)      	# Note: 1 
d2 <- d                            	# Note: 2 
d$x <- 5                           	# Note: 3 
        
print(d)
#   x y
# 1 5 2

print(d2)
#   x y
# 1 1 2

# Note 1: 
#   Create some example data and refer to it by the name d. 

# Note 2: 
#   Create an additional reference d2 to the same data. 

# Note 3: 
#   Alter the value referred to by d. 



```




00011_informalexample_2.9_of_section_2.1.2.R


```{r 00011_informalexample_2.9_of_section_2.1.2.R }
# informalexample 2.9 of section 2.1.2 
# (informalexample 2.9 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

data <- data.frame(revenue = c(2, 1, 2),    	# Note: 1 
                   sort_key = c("b", "c", "a"), 
                   stringsAsFactors = FALSE)
print(data)

#   revenue sort_key
# 1       2        b
# 2       1        c
# 3       2        a

. <- data                                     	# Note: 2 
. <- .[order(.$sort_key), , drop = FALSE]     	# Note: 3 
.$ordered_sum_revenue <- cumsum(.$revenue)
.$fraction_revenue_seen <- .$ordered_sum_revenue/sum(.$revenue)
result <- .                                   	# Note: 4 

print(result)

#   revenue sort_key ordered_sum_revenue fraction_revenue_seen
# 3       2        a                   2                   0.4
# 1       2        b                   4                   0.8
# 2       1        c                   5                   1.0

# Note 1: 
#   Our notional, or example data. 

# Note 2: 
#   Assign our data to a temporary variable named “.”. The original values will remain available 
#   in the “data” variable, making it easy to re-start 
#   the calculation from the beginning if 
#   necessary. 

# Note 3: 
#   Use the order command to sort the rows. The “drop = FALSE” is not strictly needed, but it is 
#   good to get in the habit of including it. For 
#   single column data.frames without the “drop = 
#   FALSE” argument, the “[,]” indexing operator will 
#   convert the result to a vector, which is almost 
#   never the R user's true intent. The “drop = FALSE” 
#   argument turns off this conversion, and it is a 
#   good idea to include it “just in case” and a 
#   definite requirement when either the data.frame 
#   has a single column or when we don’t know if the 
#   data.frame has more than one column (as the 
#   data.frame comes from somewhere else). 

# Note 4: 
#   Assign the result away from “.” to a more memorable variable name. 



```




00012_informalexample_2.10_of_section_2.1.2.R


```{r 00012_informalexample_2.10_of_section_2.1.2.R }
# informalexample 2.10 of section 2.1.2 
# (informalexample 2.10 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

library("dplyr")

result <- data %>%
  arrange(., sort_key) %>%
  mutate(., ordered_sum_revenue = cumsum(revenue)) %>%
  mutate(., fraction_revenue_seen = ordered_sum_revenue/sum(revenue))



```




00013_informalexample_2.11_of_section_2.1.2.R


```{r 00013_informalexample_2.11_of_section_2.1.2.R }
# informalexample 2.11 of section 2.1.2 
# (informalexample 2.11 of section 2.1.2)  : Starting with R and data : Starting with R : R programming 

d <- data.frame(col1 = c(1, 2, 3), col2 = c(-1, 0, 1))
d$col3 <- d$col1 + d$col2
print(d)
#   col1 col2 col3
# 1    1   -1    0
# 2    2    0    2
# 3    3    1    4



```




00015_example_2.1_of_section_2.2.1.R


```{r 00015_example_2.1_of_section_2.2.1.R }
# example 2.1 of section 2.2.1 
# (example 2.1 of section 2.2.1)  : Starting with R and data : Working with data from files : Working with well-structured data from files or URLs 
# Title: Reading the UCI car data 

uciCar <- read.table(               	# Note: 1 
   '../UCICar/car.data.csv',                  	# Note: 2 
   sep = ',',                       	# Note: 3 
   header = TRUE,                   	# Note: 4 
   stringsAsFactor = TRUE           	# Note: 5 
   )
                                        
View(uciCar)                        	# Note: 6

# Note 1: 
#   Command to read from a file or URL and store the result in a new data frame object 
#   called 
#   uciCar. 

# Note 2: 
#   Filename or URL to get the data from. 

# Note 3: 
#   Specify the column or field separator as a 
#   comma. 

# Note 4: 
#   Tell R to expect a header line that defines 
#   the data column names. 

# Note 5: 
#   Tell R to convert string values to factors. This is the default behavior, so we are just using this argument to document intent. 

# Note 6: 
#   Examine the data with R’s built-in 
#   table viewer. 



```




00016_example_2.2_of_section_2.2.1.R


```{r 00016_example_2.2_of_section_2.2.1.R }
# example 2.2 of section 2.2.1 
# (example 2.2 of section 2.2.1)  : Starting with R and data : Working with data from files : Working with well-structured data from files or URLs 
# Title: Exploring the car data 

class(uciCar)
## [1] "data.frame"                     	# Note: 1 
summary(uciCar)
##    buying      maint       doors
##  high :432   high :432   2    :432
##  low  :432   low  :432   3    :432
##  med  :432   med  :432   4    :432
##  vhigh:432   vhigh:432   5more:432
##
##  persons     lug_boot    safety
##  2   :576   big  :576   high:576
##  4   :576   med  :576   low :576
##  more:576   small:576   med :576
##
##    rating
##  acc  : 384
##  good :  69
##  unacc:1210
##  vgood:  65

dim(uciCar)
## [1] 1728    7                       	# Note: 2

# Note 1: 
#   The loaded object uciCar is of type data.frame. 

# Note 2: 
#   The “[1]” is just an output sequence 
#   marker. The actual information is this: uciCar has 
#   1728 rows and 7 columns. Always try to confirm you 
#   got a good parse by at least checking that the 
#   number of rows is exactly one fewer than the 
#   number of lines of text in the original file. The 
#   difference of one is because the column header 
#   counts as a line of text, but not as a data row. 



```




00018_example_2.3_of_section_2.2.2.R


```{r 00018_example_2.3_of_section_2.2.2.R }
# example 2.3 of section 2.2.2 
# (example 2.3 of section 2.2.2)  : Starting with R and data : Working with data from files : Using R with less-structured data 
# Title: Loading the credit dataset 

setwd("../../PDSwR2/Statlog")                 	# Note: 1 
d <- read.table('../Statlog/german.data', sep=' ',
   stringsAsFactors = FALSE, header = FALSE)

# Note 1: 
#   Replace this path with the actual path where you have saved PDSwR2. 



```




00019_example_2.4_of_section_2.2.2.R


```{r 00019_example_2.4_of_section_2.2.2.R }
# example 2.4 of section 2.2.2 
# (example 2.4 of section 2.2.2)  : Starting with R and data : Working with data from files : Using R with less-structured data 
# Title: Setting column names 

d <- read.table('../Statlog/german.data',
                sep  =  " ",
                stringsAsFactors  =  FALSE, header  =  FALSE)

colnames(d) <- c('Status_of_existing_checking_account', 'Duration_in_month',
                 'Credit_history', 'Purpose', 'Credit_amount', 'Savings_account_bonds', 
                 'Present_employment_since',
                 'Installment_rate_in_percentage_of_disposable_income',
                 'Personal_status_and_sex', 'Other_debtors_guarantors',
                 'Present_residence_since', 'Property', 'Age_in_years',
                 'Other_installment_plans', 'Housing',
                 'Number_of_existing_credits_at_this_bank', 'Job',
                 'Number_of_people_being_liable_to_provide_maintenance_for',
                 'Telephone', 'foreign_worker', 'Good_Loan')
str(d)
## 'data.frame':    1000 obs. of  21 variables:
##  $ Status_of_existing_checking_account                     : chr  "A11" "A12" "A14" "A11" ...
##  $ Duration_in_month                                       : int  6 48 12 42 24 36 24 36 12 30 ...
##  $ Credit_history                                          : chr  "A34" "A32" "A34" "A32" ...
##  $ Purpose                                                 : chr  "A43" "A43" "A46" "A42" ...
##  $ Credit_amount                                           : int  1169 5951 2096 7882 4870 9055 2835 6948 3059 5234 ...
##  $ Savings_account_bonds                                   : chr  "A65" "A61" "A61" "A61" ...
##  $ Present_employment_since                                : chr  "A75" "A73" "A74" "A74" ...
##  $ Installment_rate_in_percentage_of_disposable_income     : int  4 2 2 2 3 2 3 2 2 4 ...
##  $ Personal_status_and_sex                                 : chr  "A93" "A92" "A93" "A93" ...
##  $ Other_debtors_guarantors                                : chr  "A101" "A101" "A101" "A103" ...
##  $ Present_residence_since                                 : int  4 2 3 4 4 4 4 2 4 2 ...
##  $ Property                                                : chr  "A121" "A121" "A121" "A122" ...
##  $ Age_in_years                                            : int  67 22 49 45 53 35 53 35 61 28 ...
##  $ Other_installment_plans                                 : chr  "A143" "A143" "A143" "A143" ...
##  $ Housing                                                 : chr  "A152" "A152" "A152" "A153" ...
##  $ Number_of_existing_credits_at_this_bank                 : int  2 1 1 1 2 1 1 1 1 2 ...
##  $ Job                                                     : chr  "A173" "A173" "A172" "A173" ...
##  $ Number_of_people_being_liable_to_provide_maintenance_for: int  1 1 2 2 2 2 1 1 1 1 ...
##  $ Telephone                                               : chr  "A192" "A191" "A191" "A191" ...
##  $ foreign_worker                                          : chr  "A201" "A201" "A201" "A201" ...
##  $ Good_Loan                                               : int  1 2 1 1 2 1 1 1 1 2 ...



```




00021_example_2.5_of_section_2.2.2.R


```{r 00021_example_2.5_of_section_2.2.2.R }
# example 2.5 of section 2.2.2 
# (example 2.5 of section 2.2.2)  : Starting with R and data : Working with data from files : Using R with less-structured data 
# Title: Transforming the car data 

source("../Statlog/mapping.R")                      	# Note: 1 
for(ci in colnames(d)) {                     	# Note: 2 
   if(is.character(d[[ci]])) {
      d[[ci]] <- as.factor(mapping[d[[ci]]])  	# Note: 3 
   }
}

# Note 1: 
#   This file can be found at https://github.com/WinVector/PDSwR2/blob/master/Statlog/../Statlog/mapping.R . 

# Note 2: 
#   Prefer using column names to column indices. 

# Note 3: 
#   The [[]] notation is using the fact 
#   that data.frames are named lists of columns. So we 
#   are working on each column in turn. Notice the 
#   mapping lookup is vectorized: i.e. it is applied 
#   to all elements in the column in one step. 



```




00022_example_2.6_of_section_2.2.2.R


```{r 00022_example_2.6_of_section_2.2.2.R }
# example 2.6 of section 2.2.2 
# (example 2.6 of section 2.2.2)  : Starting with R and data : Working with data from files : Using R with less-structured data 
# Title: Summary of Good_Loan and Purpose 

setwd("../../PDSwR2/Statlog")               	# Note: 1 
d <- readRDS("creditdata.RDS")        	# Note: 2 

table(d$Purpose, d$Good_Loan) 
                     
##                       BadLoan GoodLoan
##   business                 34       63
##   car (new)                89      145
##   car (used)               17       86
##   domestic appliances       4        8
##   education                22       28
##   furniture/equipment      58      123
##   others                    5        7
##   radio/television         62      218
##   repairs                   8       14
##   retraining                1        8

# Note 1: 
#   Set the working directory. You will have to replace ../../PDSwR2/Statlog with the actual full path to Statlog on your machine. 

# Note 2: 
#   Read the prepared statlog data. 



```




00024_example_2.8_of_section_2.3.1.R


```{r 00024_example_2.8_of_section_2.3.1.R }
# example 2.8 of section 2.3.1 
# (example 2.8 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 
# Title: Loading data into R from a relational database 

library("DBI")
library("dplyr")                                    	# Note: 1 
library("rquery")

dlist <- readRDS("../PUMS/PUMSsample.RDS")                   	# Note: 2 
db <- dbConnect(RSQLite::SQLite(), ":memory:")      	# Note: 3 
dbWriteTable(db, "dpus", as.data.frame(dlist$ss16pus)) 	# Note: 4 
dbWriteTable(db, "dhus", as.data.frame(dlist$ss16hus))
rm(list = "dlist")                                  	# Note: 5 

dbGetQuery(db, "SELECT * FROM dpus LIMIT 5")            	# Note: 6 

dpus <- tbl(db, "dpus")                         	# Note: 7 
dhus <- tbl(db, "dhus")

print(dpus)                                              	# Note: 8 
glimpse(dpus)

View(rsummary(db, "dpus"))                     	# Note: 9

# Note 1: 
#   Attach some packages we wish to use commands 
#   and functions from. 

# Note 2: 
#   Load the data from the compressed RDS disk 
#   format into R memory. Note: you will need to 
#   change the path PUMSsample to where you have saved 
#   the contents of PDSwR2/PUMS. 

# Note 3: 
#   Connect to a new RSQLite in-memory database. 
#   We are going to use RSQLite for our examples. In 
#   practice you would connect to a pre-existing 
#   database such as PostgreSQL or Spark with 
#   pre-existing tables. 

# Note 4: 
#   Copy the data from the in-memory structure 
#   dlist into the database. 

# Note 5: 
#   Remove our local copy of the data, as we are simulating having found the data in the database. 

# Note 6: 
#   Use the SQL query language for a quick look at up to 5 rows of our data. 

# Note 7: 
#   Build dplyr handles that refer to the remote database data. 

# Note 8: 
#   Use dplyr to examine and work with the remote data. 

# Note 9: 
#   Use the rquery package to get a summary of the remote data. 



```




00025_example_2.9_of_section_2.3.1.R


```{r 00025_example_2.9_of_section_2.3.1.R }
# example 2.9 of section 2.3.1 
# (example 2.9 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 
# Title: Loading data from a database 

dpus <- dbReadTable(db, "dpus")                                  	# Note: 1 

dpus <- dpus[, c("AGEP", "COW", "ESR",  "PERNP", 
                 "PINCP","SCHL", "SEX", "WKHP")]   	# Note: 2 
                 
for(ci in c("AGEP", "PERNP", "PINCP", "WKHP")) {   	# Note: 3 
  dpus[[ci]] <- as.numeric(dpus[[ci]])
}

dpus$COW <- strtrim(dpus$COW, 50)                                  	# Note: 4 

str(dpus)                                                       	# Note: 5

# Note 1: 
#   Copy data from the database into R memory. This assumes we are continuing from the previous 
#   example, so the packages we have attached are 
#   still available and the database handle 
#   db is still valid. 

# Note 2: 
#   Select a subset of columns we want to work with. Restricting columns is not required, but 
#   improves legibility of later printing. 

# Note 3: 
#   All of the columns in this copy of PUMS data are stored as character type to preserve features such as leading zeros 
#   from the original data. Here we are converting columns we wish to treat as numeric to the numeric type. Non-numeric values, often missing entries, 
#   get coded with the symbol NA which stands for not available. 

# Note 4: 
#   The PUMS level names are very long (which is one of the reasons these columns are distributed 
#   as integers), so for this dataset that has level 
#   names instead of level codes we are shortening the 
#   employment codes to no more than 50 
#   characters. 

# Note 5: 
#   Look at the first few rows of data in a column orientation. 



```




00026_example_2.10_of_section_2.3.1.R


```{r 00026_example_2.10_of_section_2.3.1.R }
# example 2.10 of section 2.3.1 
# (example 2.10 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 
# Title: Re-mapping values and selecting rows from data 

target_emp_levs <- c(                                          	# Note: 1 
  "Employee of a private for-profit company or busine",
  "Employee of a private not-for-profit, tax-exempt, ",
  "Federal government employee",                    
  "Local government employee (city, county, etc.)",   
  "Self-employed in own incorporated business, profes",
  "Self-employed in own not incorporated business, pr",
  "State government employee")


complete <- complete.cases(dpus)                              	# Note: 2 

stdworker <- with(dpus,                                        	# Note: 3 
                  (PINCP>1000) & 
                    (ESR=="Civilian employed, at work") & 
                    (PINCP<=250000) & 
                    (PERNP>1000) & (PERNP<=250000) & 
                    (WKHP>=30) & 
                    (AGEP>=18) & (AGEP<=65) & 
                    (COW %in% target_emp_levs))
                    
dpus <- dpus[complete & stdworker, , drop = FALSE]              	# Note: 4 

no_advanced_degree <- is.na(dpus$SCHL) |                     	# Note: 5 
  (!(dpus$SCHL %in% c("Associate's degree",
                      "Bachelor's degree",
                      "Doctorate degree",
                      "Master's degree",
                      "Professional degree beyond a bachelor's degree")))
dpus$SCHL[no_advanced_degree] <- "No Advanced Degree"

dpus$SCHL <- relevel(factor(dpus$SCHL),                     	# Note: 6 
                     "No Advanced Degree")                
dpus$COW <- relevel(factor(dpus$COW), 
                    target_emp_levs[[1]])
dpus$ESR <- relevel(factor(dpus$ESR), 
                    "Civilian employed, at work")
dpus$SEX <- relevel(factor(dpus$SEX), 
                    "Male")
                    
saveRDS(dpus, "dpus_std_employee.RDS")                           	# Note: 7 
                    
summary(dpus)                                                  	# Note: 8

# Note 1: 
#   Define a vector of employment definitions we consider “standard.” 

# Note 2: 
#   Build a new logical vector indicating which rows have valid values in all of our columns of 
#   interest. In real applications, dealing with 
#   missing values is important and cannot always be 
#   handled by skipping incomplete rows. We will 
#   return to the issue of properly dealing with 
#   missing values when we discuss managing 
#   data. 

# Note 3: 
#   Build a new logical vector indicating which workers we consider typical full-time employees. 
#   Each of these column names are the ones we 
#   discussed earlier. The results of any analysis 
#   will be heavily influenced by this definition, so 
#   in a real task we would spend a lot of time 
#   researching the choices in this step. It literally 
#   controls who and what we are studying. Notice to 
#   keep things simple and homogeneous we restricted 
#   this study to civilians, which would be an 
#   unacceptable limitation in a complete work. 

# Note 4: 
#   Restrict only to rows or examples that meet or definition of typical worker. 

# Note 5: 
#   Re-code education, merging the less-than-bachelor’s-degree levels to the single level “No 
#   Advanced Degree”. 

# Note 6: 
#   Convert our string-valued columns to factors, picking the reference level with the re-level() 
#   function. 

# Note 7: 
#   Save this data to a file so we can use it in later examples. This file is also already 
#   available at the path 
#   PDSwR2/PUMS/dpus_std_employee.RDS 

# Note 8: 
#   Take a look at our data. One of the advantages of factors is summary() builds up useful counts 
#   for them. However, it was best to delay having 
#   string codes as factors until after we finished 
#   with re-mapping level codes. 



```




00027_informalexample_2.15_of_section_2.3.1.R


```{r 00027_informalexample_2.15_of_section_2.3.1.R }
# informalexample 2.15 of section 2.3.1 
# (informalexample 2.15 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 

levels(dpus$SCHL)                                 	# Note: 1 
## [1] "No Advanced Degree"                             "Associate's degree"                            
## [3] "Bachelor's degree"                              "Doctorate degree"                              
## [5] "Master's degree"                                "Professional degree beyond a bachelor's degree"

head(dpus$SCHL)                                  	# Note: 2 
## [1] Associate's degree Associate's degree Associate's degree No Advanced Degree Doctorate degree   Associate's degree
##   6 Levels: No Advanced Degree Associate's degree Bachelor's degree Doctorate degree ... Professional degree beyond a bachelor's degree

str(dpus$SCHL)                                   	# Note: 3 
##  Factor w/ 6 levels "No Advanced Degree",..: 2 2 2 1 4 2 1 5 1 1 ...

# Note 1: 
#   Show the possible levels for SCHL. 

# Note 2: 
#   Show the first few string values for SCHL. 

# Note 3: 
#   Show how the first few levels are represented as codes. 



```




00028_informalexample_2.16_of_section_2.3.1.R


```{r 00028_informalexample_2.16_of_section_2.3.1.R }
# informalexample 2.16 of section 2.3.1 
# (informalexample 2.16 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 

d <- cbind(                                                        	# Note: 1 
  data.frame(SCHL = as.character(dpus$SCHL),                       	# Note: 2 
             stringsAsFactors = FALSE),
  model.matrix(~SCHL, dpus)                                        	# Note: 3 
)
d$'(Intercept)' <- NULL                                        	# Note: 4 
str(d)                                                             	# Note: 5 

## 'data.frame':    41305 obs. of  6 variables:
##  $ SCHL                                              : chr  "Associate's degree" "Associate's degree" "Associate's degree" "No Advanced Degree" ...
##  $ SCHLAssociate's degree                            : num  1 1 1 0 0 1 0 0 0 0 ...
##  $ SCHLBachelor's degree                             : num  0 0 0 0 0 0 0 0 0 0 ...
##  $ SCHLDoctorate degree                              : num  0 0 0 0 1 0 0 0 0 0 ...
##  $ SCHLMaster's degree                               : num  0 0 0 0 0 0 0 1 0 0 ...
##  $ SCHLProfessional degree beyond a bachelor's degree: num  0 0 0 0 0 0 0 0 0 0 ...

# Note 1: 
#   The cbind operator combines two data frames by columns, or each row is built by matching columns from rows in each data frame. 

# Note 2: 
#   Build a data.frame with the SCHL column re-coded as character strings instead of as a factor. 

# Note 3: 
#   Build a matrix with dummy variables generated from the SCHL factor column. 

# Note 4: 
#   Remove a column named "(Intercept)" from the data.frame, as it is a side-effect of model.matrix we are not interested in at this time. 

# Note 5: 
#   Show the structure that presents the original SCHL string form along with the indicators. 
#   str() presents the first few rows in transpose format (flipped so rows are now up and down and columns are across). 



```




00029_informalexample_2.17_of_section_2.3.1.R


```{r 00029_informalexample_2.17_of_section_2.3.1.R }
# informalexample 2.17 of section 2.3.1 
# (informalexample 2.17 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 

table(schooling = dpus$SCHL, sex = dpus$SEX)                 	# Note: 1 
                                                
##                                                 sex
## schooling                                         Male Female
##   No Advanced Degree                             13178   9350
##   Associate's degree                              1796   2088
##   Bachelor's degree                               4927   4519
##   Doctorate degree                                 361    269
##   Master's degree                                 1792   2225
##   Professional degree beyond a bachelor's degree   421    379

tapply(                                                   	# Note: 2 
   dpus$PINCP,                                       	# Note: 3 
   list(dpus$SCHL, dpus$SEX),                        	# Note: 4 
   FUN = mean                                        	# Note: 5 
   )

##                                                     Male   Female
## No Advanced Degree                              44304.21 33117.37
## Associate's degree                              56971.93 42002.06
## Bachelor's degree                               76111.84 57260.44
## Doctorate degree                               104943.33 89336.99
## Master's degree                                 94663.41 69104.54
## Professional degree beyond a bachelor's degree 111047.26 92071.56

# Note 1: 
#   Use the table command to count how even each pair of SCHL and SEX occurs. 

# Note 2: 
#   Use tapply to tally how often each pair of SCHL of SEX occurs 

# Note 3: 
#   This argument is the vector of data we are aggregating or summarizing in the tapply. 

# Note 4: 
#   This argument list specifies how we are grouping the data, in this case simultaneously by SCHL and SEX. 

# Note 5: 
#   The argument specifies how we are aggregating values, in this case we are taking the mean or average using the mean function. 



```




00030_informalexample_2.18_of_section_2.3.1.R


```{r 00030_informalexample_2.18_of_section_2.3.1.R }
# informalexample 2.18 of section 2.3.1 
# (informalexample 2.18 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 

library("dplyr")

dpus %>%
  group_by(., SCHL, SEX)  %>%
  summarize(.,
            count = n(),
            mean_income = mean(PINCP)) %>%
  ungroup(.) %>%
  arrange(., SCHL, SEX)

## # A tibble: 12 x 4
##    SCHL                                           SEX    count mean_income
##    <fct>                                          <fct>  <int>       <dbl>
##  1 No Advanced Degree                             Male   13178      44304.
##  2 No Advanced Degree                             Female  9350      33117.
##  3 Associate's degree                             Male    1796      56972.
##  4 Associate's degree                             Female  2088      42002.
##  5 Bachelor's degree                              Male    4927      76112.
##  6 Bachelor's degree                              Female  4519      57260.
##  7 Doctorate degree                               Male     361     104943.
##  8 Doctorate degree                               Female   269      89337.
##  9 Master's degree                                Male    1792      94663.
## 10 Master's degree                                Female  2225      69105.
## 11 Professional degree beyond a bachelor's degree Male     421     111047.
## 12 Professional degree beyond a bachelor's degree Female   379      92072.



```




00031_example_2.11_of_section_2.3.1.R


```{r 00031_example_2.11_of_section_2.3.1.R }
# example 2.11 of section 2.3.1 
# (example 2.11 of section 2.3.1)  : Starting with R and data : Working with relational databases : A production-size example 
# Title: Plotting the data 

WVPlots::ScatterHist(
  dpus, "AGEP", "PINCP",
  "Expected income (PINCP) as function age (AGEP)",
  smoothmethod = "lm",
  point_alpha = 0.025)



```


